---
title: "Test boot_loess"
author: "Craig Slinkman"
date: "12/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective to test the fit_loess function

We first prepare the environment for the test. 

```{r prepare_envronment}
library( tidyverse )                    # I live in the tidyverse R-environment.
```

## Specify source for fit_loess

```{r}
source('D:/R-Projects/RStatistics/fit_loess_v01.R')

```


## Prepare data

We will be using the base R provided data set ##faithful##.  

I don't like the variable names I will change them.  I will also create a tibble names old_faithful. I do this because I think that variable names for statistical computing should be as descrptive as possible.

```{r prepare_test_data}
data( faithful  )                       # Prepare data

waiting_time    <- faithful$waiting
eruption_length <- faithful$eruptions

old_faithful <- 
  tibble( eruption_length,
          waiting_time )

old_faithful
```

Note that all times are in minutes.  


## Plot the data

The first sep in statistical analysis is to plot the data.  We draw a scatterplot below.

```{r plot_data}

ggplot( old_faithful,
        aes( x = eruption_length,
             y = waiting_time ) ) +
  geom_point() +
  xlab( "Length of prior erupotion (minutes)" ) +
  ylab( "Waiting time to next eruption (minutes)" ) +
  ggtitle( "Waiting time to next eruption versus length of prior eruption" ) +
  theme_bw()
```

## Test: fit_loess()

```{r test_fit_loess}
results <-                                        # Fit the loess.
  fit_loess( waiting_time ~ eruption_length,
             old_faithful )

smooth   <- results$smooth 
residual <- results$residual

results <- tibble( eruption_length,
                   waiting_time,
                   smooth,
                   residual )

results

ggplot( results, 
        aes( x= eruption_length, y = waiting_time ))+
  geom_point() +
  geom_line( aes( x = eruption_length, y = smooth), color = "red") +
   xlab( "Length of prior erupotion (minutes)" ) +
  ylab( "Waiting time to next eruption (minutes)" ) +
  ggtitle( "Waiting time to next eruption versus length of prior eruption with loess smooth" ) +
  theme_bw()

```

## Computation of pseedo-coeffint of determination $R^2$

```{r}
n   <- dim(results)[1]
SSY <- (n - 1) *sd (results$waiting_time )^2
SSE <- sum( results$residual^2 )
R2 <- 1 - SSE/SSY
R2


r <- cor( results$waiting_time, results$smooth )
R2 <- r^2
R2
```

